---
title: "General Overview of Rockfish Data"
output: 
  html_notebook:
    toc: true
---


This document explores the data and meta-data we have to make sure that everything
is looking like it should, and to do a few summaries.

To start off with, let's load data and libs:
```{r load-stuff}
library(tidyverse)

genos <- readRDS("../extdata/processed/called_genos_na_explicit.rds")
samples <- readRDS("../data/processed/sample-sheet-tibble.rds")
meta <- readRDS("../data/processed/meta-data-tibble.rds")
```

## Whom have we not genotyped?

Let us start off by seeing whom in the meta data was not genotyped:
```{r not-genoed}
not_geno <- meta %>%
  filter(!(NMFS_DNA_ID %in% samples$NMFS_DNA_ID))

# summarize by species and reported life stage:
not_geno %>% 
  group_by(GENUS, SPECIES, REPORTED_LIFE_STAGE) %>%
  tally() %>% 
  arrange(desc(n))
```
That looks very right.  The total number of individuals in the meta data that were
not genotyped is `r nrow(not_geno)`, and they are all in "non-target" species, or were
some of the extruded larvae that were not genotyped.  

## Whom _have_ we genotyped?

Same game with the ones that we have genotyped (or at least tried to genotype)
```{r yes-geno}
tried <- meta %>%
  filter(NMFS_DNA_ID %in% samples$NMFS_DNA_ID) 

tried %>%
  group_by(GENUS, SPECIES, REPORTED_LIFE_STAGE) %>%
  tally() %>% 
  arrange(desc(n))
```
OK, that looks right-ish.  That is `r nrow(tried)` NMFS_DNA_IDs that were put on the sequencer.  

## Is anyone missing from the metadata file?
Here are all the 
NMFS_DNA_IDs that appear in the sample_sheets but do not appear in the metadata: (or at least were not
read correctly out of the metadata file.)
```{r missing-from-meta}
missing_from_meta <- samples %>%
  filter(!(NMFS_DNA_ID %in% meta$NMFS_DNA_ID))
missing_from_meta
```
There are `r nrow(missing_from_meta)` of those.  Here is a breakdown of them:
```{r break-it-down}
missing_from_meta %>%
  group_by(gtseq_run) %>% 
  tally()
```
So, I don't know what is wrong with that single individual on GTSeq Run 47, but I am not
going to worry about him for now.  The 65 individuals from 
Run 54 are from Hailey's Species ID project (they were filling out a MiSeq lane), so it is correct
that we have no meta data for them.



