---
title: "Doing Parentage on the Kelp Rockfish"
output: 
  html_notebook:
    toc: true
---


This is a first stab at winnowing down our samples to just a few species of rockfish,
and figuring out which they are.  I am going to go for this with structure.


To start off with, let's load data and libs:
```{r load-stuff}
library(tidyverse)
library(CKMRsim)
library(stringr)

meta <- readRDS("../data/processed/meta-data-tibble.rds")
genos <- readRDS("../extdata/processed/genos-aggregated-and-no-hi-missers.rds")
species <- readRDS("../data/processed/juvie-gsi-sim-assignments.rds") 
```


## Selecting our fish and designating allele nomenclature

We gotta pick out the _S. atrovirens_.  For adults that is pretty much already done, just by who they
are, but we must use the GSI assignments for the juveniles.
```{r define-kelps}
kelp_ad <- meta %>% 
  filter(SPECIES == "atrovirens",
         REPORTED_LIFE_STAGE == "ADULT") %>%
  select(NMFS_DNA_ID) %>% unlist() %>% unname()

kelp_juv <- species %>%
  filter(score > 99.9) %>%
  select(NMFS_DNA_ID) %>% unlist() %>% unname

```

With those we can just filter down the genos to the ones that we want, and then we
can get it into the format required for CKMR.
```{r kelp-genos}
kelp_genos <- genos %>%
  filter(NMFS_DNA_ID %in% c(kelp_ad, kelp_juv))

# we will use this some more
kg2 <- kelp_genos %>% 
  select(NMFS_DNA_ID, locus, allele) %>%
  mutate(Chrom = "GTseq") %>% 
  mutate(Pos = as.integer(factor(locus, levels = unique(locus)))) %>%
  rename(Locus = locus,
         Allele = allele) %>%
  select(NMFS_DNA_ID, Chrom, Locus, Pos, Allele) %>%
  ungroup()

# get the allele freqs
kg_ckmr_markers <- kg2 %>%
  filter(!is.na(Allele)) %>% # it is vital to filter out the NAs at this stage
  group_by(Chrom, Locus, Pos, Allele) %>%
  summarise(counts = n()) %>%
  group_by(Locus, Pos) %>%
  mutate(Freq = counts / sum(counts)) %>%
  select(-counts) %>%
  mutate(AlleIdx = 1,
         LocIdx = 1) %>%
  reindex_markers(.)
```
Great! We have a total `r nrow(kg_ckmr_markers)` alleles in this data set.


## Doing CKMR simulations to assess power

Really simply let's just do this to see what our power for parentage ought to look like.
```{r CKMR-create}
CK <- create_ckmr(kg_ckmr_markers, kappa_matrix = kappas[c("PO", "FS", "U"), ])

# see what it tells us when we print it:
CK
```

Then do the simulations to see what the false positive rates are for finding parents when the truth
is U or FS:
```{r CKMR-sim-and-sample}
Qs <- simulate_Qij(C = CK, froms = c("PO", "FS", "U"), tos = c("PO", "FS", "U"), reps = 10^4)

# then do the importance sampling to get the FPRs
mc_sample_simple(Qs, nu = "PO", de = c("U", "FS"), tr = c("U", "FS"), method = "both")
```
We can come back to that later to see about finding Full Sibs.

## Making genotype matrices
```{r}
kelp_haps <- kelp_genos %>%
  filter(!is.na(allele)) %>%  # once again, it is critical to remove these at this point
  select(NMFS_DNA_ID, locus, gene_copy, allele) %>%
  rename(Locus = locus, Allele = allele)

kelp_idx_frame <- kg_ckmr_markers %>%
  select(Locus, Allele, LocIdx, AlleIdx) %>%
  group_by(Locus) %>%
  mutate(NumA = n()) %>%  # get the number of alleles at each locus
  ungroup() %>%
  left_join(kelp_haps, .)  %>%  # join the alle_idx's onto the actual genotype data
  select(NMFS_DNA_ID, Locus, gene_copy, LocIdx, NumA, AlleIdx) %>%
  spread(key = gene_copy, value = AlleIdx) %>%
  mutate(GenoIdx = index_ab(a = `1`, b = `2`, A = NumA))

# make a matrix of genotype integers 
wide_kelp <- kelp_idx_frame %>%
  select(NMFS_DNA_ID, LocIdx, GenoIdx) %>%
  spread(data = ., key = LocIdx, value = GenoIdx)
```

Don't forget to set NA's to 0, and then decrease each value by 1:
```{r make-mat}
rocky_mat <- as.matrix(wide_kelp[, -1])
rownames(rocky_mat) <- wide_kelp$NMFS_DNA_ID
rocky_mat[is.na(rocky_mat)] <- 0
rocky_mat <- rocky_mat - 1
storage.mode(rocky_mat) <-  "integer"
```

## Looking for duplicated samples

We can quickly look through rocky_mat for pairs of indivs with lots of matching genotypes.
```{r check-for-dupes}
matchers <- pairwise_geno_id(S = rocky_mat, max_miss = 12) %>%
  arrange(num_mismatch)
matchers
```
We need to deal with these somehow.  I am going to say that anyone with 3 or fewer mismatches is certainly 
the same individual, and we are going to take just one from amongst those pairs/groups.  We use a hasty method:
we toss out all the individuals that appear in both ind1 and ind2, and then all the remaining ind2's.  This 
might throw out one more than we need to, but it is what we do for now.
```{r toss-matchers}
def_same <- matchers %>%
  filter(num_mismatch <= 3)

toss_these <- c(intersect(def_same$ind1, def_same$ind2), def_same$ind2) %>%
  unique()

rocky_mat_tossed <- rocky_mat[-toss_these,]
```

Now we have to break that into two matrices: one for the adults and one for the kids.
This should not be too hard.  We have to deal with the fact that not all of the adults in the meta data
had few enough missing genotypes to be included.
```{r break-stage}
kelp_ad_kept <- kelp_ad[kelp_ad %in% rownames(rocky_mat_tossed)]
kelp_juv_kept <- kelp_juv[kelp_juv %in% rownames(rocky_mat_tossed)]

adult_mat <- rocky_mat_tossed[kelp_ad_kept, ]
juvie_mat <- rocky_mat_tossed[kelp_juv_kept, ]
```

## Doing the comparisons

We need to flatten the CKMR object appropriately
```{r flatten}
# flatten those out
po_flat <- flatten_ckmr(CK, "PO")
unrel_flat <- flatten_ckmr(CK, "U")

# then compute the log-likelihoods for the parent offspring relationship
po_logl_flat <- po_flat
po_logl_flat$probs <- log(po_flat$probs / unrel_flat$probs)

```


Now we are ready to do these!
```{r do-comps}
idx <- 1:nrow(juvie_mat)
names(idx) <- idx
comps <- lapply(idx, function(i) {
    tmp <- comp_ind_pairwise(S = adult_mat, T = juvie_mat, t = i, values = po_logl_flat$probs, nGenos = po_logl_flat$nGenos, Starts = po_logl_flat$base0_locus_starts)
    #tmp[rev(top_index(tmp$value, 5)), ]  # just take the top 5 from each
    }) %>%
  dplyr::bind_rows(.id = "offspring") %>%
  dplyr::tbl_df() 

# have a look at that
result <- arrange(comps, desc(value)) %>%
  mutate(number = 1:nrow(.)) %>%
  mutate(id1 = rownames(juvie_mat)[as.integer(offspring)],
         id2 = rownames(adult_mat)[ind])

result %>% filter(number <= 1000)

```

Try plotting all those pairs, and compare them to the simulated values
```{r plot-all-pairs-density}
simlogls <- extract_logls(Q = Qs, numer = c(PO = 1), denom = c(U = 1))
ggplot(result, aes(x = value)) +
  geom_density() +
  geom_density(data = simlogls, aes(x = logl_ratio, fill = true_relat), alpha = 0.5)
```

Note that some of the individuals appear to have lots of offspring:
```{r tally-adults}
result %>% 
  filter(value > 10) %>% 
  group_by(ind, id2) %>% 
  tally() %>% 
  arrange(desc(n))
```

What if we do that with the offspring.  Hopefully we won't find mutliple parents (or at least
no more than two).
```{r tally-juveniles}
offnums <- result %>% 
  filter(value > 10) %>% 
  group_by(offspring, id1) %>% 
  tally() %>% 
  arrange(desc(n))

offnums
```

Holy Smokes! That is just what it should look like if this were correct.  That is encouraging!

## Let's take a closer look at these apparently large families

```{r checking-again}
num_offs_frame <- result %>% 
  filter(value > 10) %>% 
  group_by(ind, id2) %>% 
  mutate(num_offs = n()) %>%
  arrange(desc(num_offs), id2, desc(value))

num_offs_frame
```

## A closer look at the offspring with apparently two parents

```{r closer-offspring}
 result %>% 
  filter(value > 10) %>% 
  group_by(offspring, id1) %>% 
  filter(n() == 2) %>%
  arrange(offspring) %>%
  summarise(mapa1 = min(ind),
         mapa2 = max(ind)) %>%
  group_by(mapa1, mapa2) %>% 
  tally()
```

OK, that is kind of ugly looking.  

Hmmm....we really need to look at these on a locus-by-locus basis.



