---
title: "Running the suspicious kelp rockfish sib-like adults through Colony"
output: 
  html_notebook:
    toc: true
---


```{r load-em}
library(tidyverse)
library(CKMRsim)
```

## Introduction

So, to get to this point we had to run through "sorting-to-species.Rmd" and also
"kelp-rockfish-parentage.Rmd".  That last one creates an output file RDS that we can 
use here:
```{r load-data}
dat <- readRDS("rds_outputs/84-loci-kr-parentage-output-list.rds")
```


## Preparing Data

First get the genotype data into the format that CKMR needs to convert
to colony.  Basically just need to change some column headings
```{r format-genos}
genos <- dat$kelp_idx_frame %>%
  rename(id = NMFS_DNA_ID,
         alle1 = `1`,
         alle2 = `2`) %>%
  select(id, Locus, LocIdx, alle1, alle2)

genos
```


Now, what we need to do is pick out some 100 individuals that were not dupes, did not 
have aberrant internal HZ and were not part of the sibling pairs that
we found, and were not juvenile members of the parent-offspring pairs we found, 
and we will add those to the 19 indvividuals that were part of that
big sibship-looking thing.  
```{r find-bland}
everyone <- unique(genos$id)

# here about 270 fish we don't want in there
tossers <- c(dat$duplicate_samples$id1,
             dat$duplicate_samples$id2,
             dat$top_po_pairs$id1, 
             dat$int_hz$NMFS_DNA_ID[dat$int_hz$fract_hz > .60 | dat$int_hz$fract_hz < .27],
             unlist(dat$CnC_sort[-1]))

set.seed(5)
extras <- sample(setdiff(everyone, tossers), size = 150, replace = FALSE)

for_colony <- c(dat$CnC_sort[[1]], extras)  # about 169 individuals

genos_for_colony <- genos %>%
  filter(id %in% for_colony)
```
So, now we have 168 individuals, 19 or which were part of that sibship-looking thing.

## A weird finding

Holy Smokes! Check this out.  Of the 19 fish in the sibship-like thing, three of them also turned
up as possible parent pairs.  (Though not after dropping the null-allele-ish loci.)
```{r check-common-fish}
shared <- intersect(dat$CnC_sort[[1]], dat$top_po_pairs$id2)
dat$top_po_pairs %>%
  filter(id2 %in% shared)
```

Oh my.  Those are the three individuals that looked like they could have been parents of that one contaminated
sample (R031681 had super high internal heterozygosity).  The fact that some  these individuals that looked like
they belonged in a big sibship also have high PO logl with this contaminated indiv is worrying.  It suggests 
that maybe there is something going on with these particular individuals that is sprurious.

## Make a colony data set

```{r colony-create}
dir.create("../colony_area/84-loci-siblike-group/")
ckmr2colony(CKF = dat$kg_ckmr_markers, CKG = genos_for_colony)
file.copy("Colony2.dat", "../colony_area/84-loci-siblike-group/")
```


Then I ran that on the command line.


## Checking out the results

I can slurp the inferred pedigree out as follows.  This gets it for all 96 loci at the
moment, cuz it is still being run on the 84 loci...
```{r get-ped}
ped <- afblue::read_colony_best_config("../colony_area/siblike-group/output.BestConfig") %>%
  filter(!is.na(mom))
```

If we want to get just the individuals that were in the original 19-member siblike group
we filter it, and then we might as well arrange it to make it easier to see
```{r filter-ped}
ped %>%
  filter(id %in% dat$CnC_sort[[1]]) %>%
  arrange(dad, mom)
```


So, that makes it look like it is consistent with two half-sibships.  But that seems a little
far-fetched to me.  Are we to expect that all these fish mated amongst themselves and 
happened to have super high survival?  Maybe....In one way it would make sense if all the 
fish in one little packet of ocean happened to do well, but I still don't like it.  

It is suspicious to me that there are no full-sibships of size 3 or greater.  You can't exclude
a sib group of size 2, so there is just very little information there.  


### Let's look at the 84-locus result too

```{r ped-84}
ped84 <- afblue::read_colony_best_config("../colony_area/84-loci-siblike-group/output.BestConfig") %>%
  filter(!is.na(mom))

ped84 %>%
  filter(id %in% dat$CnC_sort[[1]]) %>%
  arrange(dad, mom)
```
Holy Cow! That is not what I was expecting to see.  Could it be that the homozygosity from the null alleles
at 12 loci was what was really making these guys unsuitable as full sibs?  Or is this just
a poor-mixing result from Colony?

I think it is time for another visual inspection of the genotype graphs
```{r geno-graphs}
col_full_sib <- ped84$id[ped84$mom == "ma_1" & ped84$dad == "pa_1"]
genos %>%
  filter(id %in% col_full_sib) %>%
  mutate(graph = ifelse(alle1<=alle2, paste(alle1, alle2, sep = "-"), paste(alle2, alle1, sep = "-"))) %>%
  group_by(Locus, LocIdx, graph) %>%
  tally()
  
```

From that it is super clear that there is not way that is a proper full-sibling group, unless there is
a boatload of genotyping error.  

What do the read depths look like for these individuals?
```{r look-at-read-depths}
haplot <- readRDS("../extdata/processed/genos-aggregated-and-no-hi-missers.rds") 

fs_read_depths <- haplot %>%
  ungroup %>%
  filter(NMFS_DNA_ID %in% col_full_sib) %>%
  filter(locus %in% unique(genos$Locus)) %>%
  arrange(locus, NMFS_DNA_ID, gene_copy) %>%
  left_join(dat$kg_ckmr_markers %>% rename(locus = Locus, allele = Allele) %>% select(locus, LocIdx, allele, AlleIdx))
  
fs_read_depths
```



